IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-07-08 paper/retrofit.tex
Fix up retrofit
Signed-off-by: Edward Z. Yang <ezyang@cs.stanford.edu>

Supporting fine-grained labeling of data in a programming language often
requires extremely intrusive modifications to its semantics and
implementation.  In this paper, we advocate applying coarse-grained
labeling~\cite{Zeldovich:2006, lio} to its extreme: our starting point
is multiple instances of completely isolated language runtimes or \emph{tasks}, with
information flow control applied between their communication.  For
languages which are implemented with isolation, such as JavaScript or C
with access to hardware privilege separation, systems along these lines
are far more easily implementable and require minimal semantic changes.

Why is this setting more interesting than ``traditional
coarse-grained information flow control system on isolated
threads?''  First, although completely separate tasks
are semantically easy to reason about, real-world implementations often blur the
lines between tasks in the name of efficiency.  Characterizing what optimizations
are permissible is a problem we address in Section~\ref{sec:concrete}, and
it is quite subtle, since removing transitions from the operational semantics
of a language can in general break non-interference.
Second, even though the underlying
semantic model involves communicating tasks, these tasks can be coordinated
together in ways that simulate traditional features in monolithic languages.
In fact, it is a very useful design tool for discovering what variants of the
features are permissible and which are not.

A very simple example is a low security task sending a high security
task a mutable reference:
\begin{align*}
    & |let i = TI (sandbox (brecvf x _ (IT (!(TI x)))))|\\
    & |in TI (send (IT i) getLabel (IT (ref true)))|
\end{align*}
For now, ignore the tags |TI| and |IT|: roughly, this code creates a new
|sandbox|ed task with identifier $i$ which waits (\textbf{blockingRecv}) for a
message, and then \textbf{send}s the task a mutable reference to |(ref
true)|.  If this operation actually shared the mutable cell, this cell
could be used to violate information flow control, by leaking
information from the high security task to the low security task.
When working with
isolated tasks which each have their own heaps, the isolation is obviously true:
the sandboxed task receives a dangling pointer.

While this is a very simple example, we can still ask: given the
obviously safe formulation with separate heaps, can this fact be useful
for showing the safety of a semantics which only has a single heap
(we answer this question positively in Section~\ref{sec:single-heap})?
Furthermore, the na\"ive
semantics violates expectations about how inter-task communication
works on a shared heap: it strongly suggests, for example, that which
types of data may be passed between tasks should be restricted (for
example, in JavaScript, one might only allow JSON objects to be passed
between tasks, rather than general object structures).

>>>>

Supporting fine-grained labeling of data in a programming language often
requires extremely intrusive modifications to its semantics and
implementation.  In this paper, we advocate applying coarse-grained
labeling~\cite{Zeldovich:2006, lio} at its most extreme: our starting point
is multiple instances of completely isolated language runtimes or \emph{tasks}, with
information flow control applied between their communication.  For
languages which are implemented with isolation, such as JavaScript or C
with access to hardware privilege separation, systems along these lines
are far more easily implementable and require fewer semantic changes.

What distinguishes this setting from previous coarse-grained IFC
systems?
%
First, even though the underlying semantic model involves communicating
tasks, these tasks can be coordinated together in ways that simulate
traditional features in monolithic languages.  In fact, simulating
features in this way is a very useful \emph{design tool} for discovering
what variants of the features are permissible and which are not.
%
Second, although completely separate tasks are semantically easy to
reason about, real-world implementations often blur the lines between
tasks in the name of efficiency (Section~\ref{sec:concrete}).
Characterizing what optimizations are permissible is subtle, since
removing transitions from the operational semantics of a language can
break non-interference.

To give a simple example, here is a low security task sending a high security
task a mutable reference:
\begin{align*}
    & |let i = TI (sandbox (brecvf x _ (IT (!(TI x)))))|\\
    & |in TI (send (IT i) l (IT (ref true)))|
\end{align*}
For now, ignore the tags |TI| and |IT|: roughly, this code creates a new
|sandbox|ed task with identifier $i$ which waits (\textbf{blockingRecv}) for a
message, and then \textbf{send}s the task a mutable reference |(ref
true)| which is labeled $l$.  If this operation actually shared the mutable cell, this cell
could be used to violate information flow control, by leaking
information from the high security task to the low security task.

This example demonstrates both of our design principles.  When working
with isolated tasks, each of which have their own heaps, the isolation
is obviously true: the sandboxed task receives a dangling pointer.  From
there, we can further ask: given the obviously safe formulation with
separate heaps, can this fact be useful for showing the safety of a
semantics which only has a single heap (we answer this question
positively in Section~\ref{sec:single-heap})?  Finally, this na\"ive
semantics violates expectations about how inter-task communication works
on a shared heap, and strongly suggests one should restrict the types of
data which may be passed between tasks (for example, in JavaScript, one
might only allow JSON objects to be passed between tasks, rather than
general object structures).
